---
title: "AlligatorDataExploration"
author: "Anthony Johnson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Thanks for checking out the johnson_alligator.pkg, an R package created for the GBIO 508 course at Southeastern Louisiana University. This package was originally designed for a fake alligator data set, this package provides a versatile set of functions for data manipulation, transformation, analysis, and visualization. Follow the tutorial below to see how it works!


Lets start by downloading the data.

```{r, eval=FALSE}
#install.packages("googlesheets4)
library(googlesheets4)
googlesheets4::gs4_deauth()
alligators <- read_sheet("https://docs.google.com/spreadsheets/d/1hNHAhO4OhXzpw6LG0WN_AZxaEUiyDhairunCXgpVLqM/edit?usp=sharing")
```


Lets download the package as well.

```{r}
#Install the package from git hub

devtools::install_github("anthonyjohnson508/johnson_alligator.pkg")


```

Go ahead an install and load these packages before we begin.

```{r}

install.packages("tidyverse")
install.packages("broom")
install.packages("knitr")


library(tidyverse) 
library(broom)
library(knitr)

```




